---
title: "Personal_project"
author: "Joseph Jaiyeola"
date: "2022-10-15"
output: html_document
---


```{r}
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(GGally)
library(gtools)
library(caret)
library(car)
library(scales)
library(lmtest)
library(ggplot2)
library(plotly)
library(ggthemes)
library(MLmetrics)
library(performance)
library(ggpubr)


library(logistf)
library(ggplot2)
library(tidyverse)
library(corrplot)
library(lattice)
library(caret)
library(MASS)
library(caTools)



library(gam)
library(tidyverse)
library(car)
library(broom)
library(DescTools)
library(ROCR)
library(lmtest)
```




```{r}
library(readr)
healthcare <- read_csv("watson_healthcare_modified.csv")
```


```{r}
str(healthcare)
```

# Check if there is any missing
```{r}
anyNA(healthcare)
```

# Check if any duplicate data

```{r}
healthcare[duplicated(healthcare),]

# no duplicated data
```


```{r}
glimpse(healthcare)

#This dataset consist of 35 features (variables) and 1,676 observations (rows data). There are 9 categorical columns and 26 numerical columns.
# Make age a categorical variable 

```

# Here are the summary of the raw dataset:

```{r}
summary(healthcare 
        %>% select_if(is.numeric))
```


# There are some variables that can be removed as they do not give useful information nor relevant to the dependent variable

```{r}
#healthcare <- healthcare %>% select(-c("Over18", "EmployeeCount", "EmployeeID", "StandardHours", "HourlyRate", "MonthlyRate", "DailyRate"))

healthcare <- healthcare [-c(22, 10, 1, 27, 13, 20, 5)]
str(healthcare)

# Now the number of columns reduced from 35 to 28. Now let’s check if there is a missing value:
```




# Here is the first 6 data.
```{r}
head(healthcare)
```


# There was a level called admin and adminstration, so I made them same level

```{r}
#table(healthcare$JobRole)

healthcare$JobRole<- car::Recode(healthcare$JobRole,
                     recodes="'Admin'='Administrative'",
                    as.factor=T)
```




# Socio Demographic Associated variable- Gender,Education Background, Age

```{r}
detach(package: MASS, unload = TRUE)

#library(conflicted)
#conflict_prefer("select", "dplyr")

d1_plot <- healthcare %>% 
  select(Gender, Attrition) %>% 
  count(Gender, Attrition) %>% 
  ggplot(aes(x=Gender, y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.2, reverse = F), size=4) +
  labs(fill = "Attrition", y="Count") + theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())


d2_plot <- healthcare %>%
  select(EducationField, Attrition) %>% 
  count(EducationField, Attrition) %>% 
  ggplot(aes(x=EducationField, y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F), size=4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(fill = "Attrition", y = "")


d3_plot <- healthcare %>% 
  select(Department, Attrition) %>% count(Department, Attrition) %>% 
  ggplot(aes(x=Department, y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  scale_x_discrete(labels=c("Cardiology","Maternity","Neurology")) +
  labs(fill = "Attrition", y = "Count")

d4_plot <- healthcare %>% 
  select(JobRole, Attrition) %>% count(JobRole, Attrition) %>% 
  ggplot(aes(x=JobRole, y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.3, reverse = F),size=3.5) +
  coord_flip() +
  theme_minimal() + 
  theme(#axis.text.x = element_text(angle = 90),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(fill = "Attrition",
       x = "Job Role", y = "Count")

d5_plot <- healthcare %>%
  select(JobLevel, Attrition) %>% count(JobLevel, Attrition) %>% 
  ggplot(aes(x=JobLevel, y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.8, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(fill = "Attrition",
       x = "Job Level", y="")

d6_plot <- healthcare %>%
  mutate(Age = as.factor(
    ifelse(Age < 20, "18-19",
        ifelse((Age >= 20) & (Age <= 25), "20-25",
          ifelse((Age >= 26) & (Age <= 30), "26-30",
            ifelse((Age >= 31) & (Age <= 35), "31-35",
              ifelse((Age >= 36) & (Age <= 40), "36-40",
                ifelse((Age >= 41) & (Age <= 45), "41-45",
                  ifelse((Age >= 46) & (Age <= 50), "46-50",
                    ifelse((Age >= 51) & (Age <= 55), "51-55", ">55"
                    )
                  )
                )
              )
            )
          )
        )
      )
    ) 
  ) %>% 
  group_by(Age, Attrition) %>% count(Age, Attrition) %>% 
  ggplot(aes(x=factor(Age, levels = c("18-19", "20-25", "26-30", "31-35", "36-40",
                                      "41-45", "46-50", "51-55", ">55")), 
             y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.3, reverse = F),size=3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90),
         panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(fill = "Attrition", x = "Age", y = "")

demography_plot <- ggarrange(d1_plot, d2_plot, d3_plot, d5_plot,
                      ncol = 2, nrow = 2, 
                      common.legend = T, 
                      legend = "bottom")
demography_plot
```

```{r}
demography_plot2 <- ggarrange(d4_plot, d6_plot,
                      ncol = 2, nrow = 1, 
                      common.legend = T, 
                      legend = "bottom")
demography_plot2
```

# Worker Experiences variables


```{r}
# 1. TotalWorkingYears
plot_exp1 <- healthcare %>% select(TotalWorkingYears, Attrition) %>% 
  mutate(binning_workingyears = as.factor(
                                ifelse(TotalWorkingYears <= 4, "0-4", 
                                  ifelse((TotalWorkingYears >= 5 & TotalWorkingYears <= 9), "5-9",
                                    ifelse((TotalWorkingYears >= 10 & TotalWorkingYears <= 14), "10-14",
                                      ifelse((TotalWorkingYears >= 15 & TotalWorkingYears <= 19), "15-19",
                                        ifelse((TotalWorkingYears >= 20 & TotalWorkingYears <= 24), "20-24",
                                          ifelse((TotalWorkingYears >= 25 & TotalWorkingYears <= 29), "25-29",
                                            ifelse((TotalWorkingYears >= 26 & TotalWorkingYears <= 29), "26-29",
                                              ifelse((TotalWorkingYears >= 30 & TotalWorkingYears <= 34), "30-34",
                                                ifelse((TotalWorkingYears >= 35 & TotalWorkingYears <= 39), "35-39", ">=40"
                                                      )
                                                    )
                                                  )
                                                )
                                              )
                                            )
                                          )
                                        )
                                      )
                                )
         ) %>% 
  count(binning_workingyears, Attrition) %>% 
  ggplot(aes(x=factor(binning_workingyears, levels = c("0-4", "5-9", "10-14", "15-19", "20-24",
                                                       "25-29", "26-29", "30-34", "35-39", ">=40")), 
             y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.3, reverse = F), size=3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Total Working Experience",
    fill = "Attrition", x = "Total Working (Years)", y = ""
    )


# 2.NumCompaniesWorked
plot_exp2 <- healthcare %>% select(NumCompaniesWorked, Attrition) %>% 
  mutate(binning = as.factor(
                                ifelse(NumCompaniesWorked <= 1, "0-1", 
                                  ifelse((NumCompaniesWorked >= 2 & NumCompaniesWorked <= 3), "2-3",
                                    ifelse((NumCompaniesWorked >= 4 & NumCompaniesWorked <= 5), "4-5",
                                      ifelse((NumCompaniesWorked >= 6 & NumCompaniesWorked <= 7), "6-7", ">7"
                                            )
                                          )
                                        )
                                      )
                                )
         ) %>% 
  count(binning, Attrition) %>% 
  ggplot(aes(x=factor(binning, levels = c("0-1", "2-3", "4-5", "6-7",">7")), 
             y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Total Company Worked",
    fill = "Attrition", x = "Total Company Worked", y = ""
    )

# 3.YearsAtCompany
plot_exp3 <- healthcare %>% select(YearsAtCompany, Attrition) %>% 
  mutate(binning_years = as.factor(
                                ifelse(YearsAtCompany <= 4, "0-4", 
                                  ifelse((YearsAtCompany >= 5 & YearsAtCompany <= 9), "5-9",
                                    ifelse((YearsAtCompany >= 10 & YearsAtCompany <= 14), "10-14",
                                      ifelse((YearsAtCompany >= 15 & YearsAtCompany <= 19), "15-19",
                                        ifelse((YearsAtCompany >= 20 & YearsAtCompany <= 24), "20-24",
                                          ifelse((YearsAtCompany >= 25 & YearsAtCompany <= 29), "25-29",
                                            ifelse((YearsAtCompany >= 26 & YearsAtCompany <= 29), "26-29",
                                              ifelse((YearsAtCompany >= 30 & YearsAtCompany <= 34), "30-34",
                                                ifelse((YearsAtCompany >= 35 & YearsAtCompany <= 39), "35-39", ">=40"
                                                      )
                                                    )
                                                  )
                                                )
                                              )
                                            )
                                          )
                                        )
                                      )
                                )
         ) %>% 
  count(binning_years, Attrition) %>% 
  ggplot(aes(x=factor(binning_years, levels = c("0-4", "5-9", "10-14", "15-19", "20-24",
                                                       "25-29", "26-29", "30-34", "35-39", ">=40")), 
             y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.2, reverse = F), size=3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Years at Company",
    fill = "Attrition", x = "Years at Company", y = ""
    )

# 4.NumCompaniesWorked
plot_exp4 <- healthcare %>% select(TrainingTimesLastYear, Attrition) %>% 
  mutate(binning = as.factor(
                                ifelse(TrainingTimesLastYear <= 1, "0-1", 
                                  ifelse((TrainingTimesLastYear >= 2 & TrainingTimesLastYear <= 3), "2-3",
                                    ifelse((TrainingTimesLastYear >= 4 & TrainingTimesLastYear <= 5), "4-5", ">5"
                                          )
                                        )
                                      )
                                )
         ) %>% 
  count(binning, Attrition) %>% 
  ggplot(aes(x=factor(binning, levels = c("0-1", "2-3", "4-5", ">5")), 
             y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.8, reverse = F), size=4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Total Completed Training",
    fill = "Attrition", x = "Total Training Completed", y = ""
    )

experience_plot <- ggarrange(plot_exp1, plot_exp2, plot_exp3, plot_exp4,
                      ncol = 2, nrow = 2, 
                      common.legend = T, 
                      legend = "bottom")
experience_plot
```




# Company Survey

```{r}
#1. EnvironmentSatisfaction
plot_sv1 <- healthcare %>% select(EnvironmentSatisfaction, Attrition) %>% 
  count(EnvironmentSatisfaction, Attrition) %>% 
  ggplot(aes(x=factor(EnvironmentSatisfaction), y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Employee Satisfaction",
    fill = "Attrition", x = "", y = ""
    )

#2. JobSatisfaction
plot_sv2 <- healthcare %>% select(JobSatisfaction, Attrition) %>% 
  count(JobSatisfaction, Attrition) %>% 
  ggplot(aes(x=factor(JobSatisfaction), y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Job Satisfaction",
    fill = "Attrition", x = "", y = ""
    )

#3. JobInvolvement
plot_sv3 <- healthcare %>% select(JobInvolvement, Attrition) %>% 
  count(JobInvolvement, Attrition) %>% 
  ggplot(aes(x=factor(JobInvolvement), y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Job Involvement",
    fill = "Attrition", x = "", y = ""
    )

#4. WorkLifeBalance
plot_sv4 <- healthcare %>% select(WorkLifeBalance, Attrition) %>% 
  count(WorkLifeBalance, Attrition) %>% 
  ggplot(aes(x=factor(WorkLifeBalance), y=n)) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=n, fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(
    title="Work Life Balance",
    fill = "Attrition", x = "", y = ""
    )

csurvey_plot <- ggarrange(plot_sv1, plot_sv2, plot_sv3, plot_sv4,
                      ncol = 2, nrow = 2, 
                      common.legend = T, 
                      legend = "bottom")
csurvey_plot
```


```{r}
healthcare %>% select(Attrition) %>% count(Attrition) %>% 
  mutate(percent=round((n/sum(n))*100,2), 
         lab_ypos = cumsum(percent) - 0.7*percent) %>% 
  ggplot(aes(x=2, y=percent, fill = factor(Attrition, levels = c("Yes", "No")))) +
  geom_bar(stat="identity", start=0) +
  coord_polar(theta = "y", start=0) +
  geom_text(aes(y = lab_ypos, 
                label = paste0(percent,' ','%')), color = "white") +
  theme_void() + theme(legend.position = "bottom") + xlim(0.5, 2.5) +
  labs(title = "Attrition Rate in Health Care ", fill = "Attrition")
```




# Department

```{r}
temp_plot1 <- healthcare %>% 
  select(Department, Attrition) %>% count(Department, Attrition) %>% 
  group_by(Department) %>% 
  mutate(percent = round((n/sum(n))*100,2)) %>% 
  ggplot(aes(x=Department, y=n,
             text=paste0('</br>Department: ', Department,
                         '</br>Attrition Status: ', Attrition,
                         '</br>Count: ', n,
                         '</br>Percentage: ', percent, ' ', '%'))) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=paste0(percent, ' ' ,'%') , fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Attrition Rate per Department", fill = "Attrition", y = "")

ggplotly(temp_plot1, tooltip="text")
```


```{r}
healthcare$Gender<- as.character(healthcare$Gender)
```


```{r}
temp_plot22 <- healthcare %>% 
  select(Gender, Attrition) %>% count(Gender, Attrition) %>% 
  group_by(Gender) %>% 
  mutate(percent = round((n/sum(n))*100,2)) %>% 
  ggplot(aes(x=factor(Gender), y=n,
             text=paste0('</br>Gender: ', Gender,
                         '</br>Attrition Status: ', Attrition,
                         '</br>Count: ', n,
                         '</br>Percentage: ', percent, ' ', '%'))) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=paste0(percent, ' ' ,'%') , fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Attrition Rate per Gender", fill = "Attrition", x = "Gender", y = "")

ggplotly(temp_plot22, tooltip="text")
```



#Job Level

```{r}
temp_plot2 <- healthcare %>% 
  select(JobLevel, Attrition) %>% count(JobLevel, Attrition) %>% 
  group_by(JobLevel) %>% 
  mutate(percent = round((n/sum(n))*100,2)) %>% 
  ggplot(aes(x=factor(JobLevel), y=n,
             text=paste0('</br>Job Level: ', JobLevel,
                         '</br>Attrition Status: ', Attrition,
                         '</br>Count: ', n,
                         '</br>Percentage: ', percent, ' ', '%'))) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=paste0(percent, ' ' ,'%') , fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Attrition Rate per Job Level", fill = "Attrition", x = "Job Level", y = "")

ggplotly(temp_plot2, tooltip="text")
```

# Around 63 % of employees were in Job Level 1 and 2. The highest attrition rate is from Job Level I (26.34 %), followed by Job Level III (14.68 %), Job Level II (9.74 %), Job Level V (7.25 %), and Job Level IV (4.72 %).


#Salary & Satisfaction Score

```{r}
avgincome_plot1 <-  healthcare %>% 
  select(JobRole, MonthlyIncome, Attrition) %>% 
  group_by(JobRole, Attrition) %>% 
  summarise(avg_monthly_income = round(mean(MonthlyIncome),2)) %>% 
  ggplot(aes(x=JobRole, y=avg_monthly_income,
             fill=factor(Attrition, levels = c("Yes","No")))) +
  geom_bar(stat = "identity", position = "dodge") + 
  geom_text(aes(x=JobRole, y=1000, 
                label = paste0("$"," ", avg_monthly_income)),
                fontface="bold", color="white", size = 2.5, nudge_y = 1500) +
  facet_wrap(~Attrition) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Job Role", y = "Average Income (dollar)",
    fill = "Attrition",
    title="Average Income\nby Job Role and Attrition Status"
  ) +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
avgincome_plot1
```

```{r}
median_incomevsjobsscore <- healthcare %>% select(MonthlyIncome, JobSatisfaction, Attrition) %>% 
  group_by(JobSatisfaction, Attrition) %>% 
  summarise(median_income = median(MonthlyIncome)) %>% 
  ggplot(aes(x=median_income, y=JobSatisfaction, 
             color=factor(Attrition, levels = c("Yes","No")))) +
  geom_point(size = 4) +
  geom_segment(aes(x = 0, xend = median_income, 
                   y = JobSatisfaction, yend = JobSatisfaction), size = 2) +
  geom_text(aes(x=1000, y=JobSatisfaction, 
                label= paste0("$ ", " ", median_income)),
            nudge_y = 0.2) + 
  facet_wrap(~Attrition) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.grid.major.x = element_line(linetype = "dashed", colour = "grey"),
        panel.grid.minor.x = element_line(linetype = "dashed", colour = "grey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(
    x = "Median Income (dollar)", y = "Job Satisfaction Score",
    color = "Attrition",
    title="Does Income affect on Job Satisfaction?\nby Attrition Status"
  )
median_incomevsjobsscore
```


```{r}
#summary(healthcare$PercentSalaryHike)

plot_subsalary1 <- healthcare %>% select(PercentSalaryHike, Attrition) %>% 
  count(PercentSalaryHike, Attrition) %>% 
  group_by(PercentSalaryHike) %>% mutate(percent = round((n/sum(n))*100,2)) %>% 
  ggplot(aes(x=factor(PercentSalaryHike), y=n,
             text=paste0('</br>Salary Hike Last Year (in percent): ', PercentSalaryHike,
                         '</br>Attrition Status: ', Attrition,
                         '</br>Count: ', n,
                         '</br>Percentage: ', percent, ' ', '%'))) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  geom_text(aes(label=paste0(percent, ' ' ,'%') , fill = factor(Attrition, levels = c("Yes", "No"))),
            angle=90, fontface="bold", color="white", size=2,
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Attrition Rate per Salary Hike Last Year\n in percent", fill = "Attrition", x = "Salary Hike (%)", y = "")

ggplotly(plot_subsalary1, tooltip="text")  
```



```{r}
healthcare %>% select(EnvironmentSatisfaction, JobRole, Attrition) %>% 
  group_by(JobRole, Attrition) %>% 
  summarize(avg_env_score = round(mean(EnvironmentSatisfaction),2)) %>% 
  ggplot(aes(x=JobRole,y=avg_env_score)) +
  geom_line(aes(group=Attrition), linetype= "twodash", size=1) +
  geom_point(aes(color=Attrition), size=3) + 
  theme_minimal() +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90),
        axis.line = element_line(colour = "grey", 
                      size = 0.7, linetype = "solid"),
        panel.grid.major.x = element_line(size = 0.5, linetype = "dashed", colour = "lightgray"),
        panel.grid.minor.x = element_line(size = 0.5, linetype = "dashed", colour = "lightgray"),
        panel.grid.major.y = element_line(size = 0.5,linetype = "dashed", colour = "lightgray")) +
  labs(
    x = "", y = "Average Score",
    fill = "Attrition",
    title = "Average Environtment Satisfaction Score\nper Job Role"
  )
```


```{r}
temp_plot4 <- healthcare %>% 
  select(Gender, MaritalStatus, Attrition) %>% count(Gender, MaritalStatus, Attrition) %>% 
  group_by(Gender, MaritalStatus) %>% 
  mutate(percent = round((n/sum(n))*100,2)) %>% 
  ggplot(aes(x=factor(MaritalStatus), y=n,
             text=paste0('</br>Gender: ', Gender,
                         '</br>Marital Status: ', MaritalStatus,
                         '</br>Attrition Status: ', Attrition,
                         '</br>Count: ', n,
                         '</br>Percentage: ', percent, ' ', '%'))) +
  geom_col(aes(fill=factor(Attrition, levels = c("Yes", "No")))) +
  facet_wrap(~Gender) +
  geom_text(aes(label=paste0(percent, ' ' ,'%') , fill = factor(Attrition, levels = c("Yes", "No"))),
            position = position_stack(vjust = 0.5, reverse = F)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Attrition Rate per Job Level", fill = "Attrition", x = "Marital Status", y = "")

ggplotly(temp_plot4, tooltip="text")
```







# Chi-Square Test for Feature Selection


```{r}
chisq.test(healthcare$BusinessTravel, healthcare$Attrition)
chisq.test(healthcare$Department, healthcare$Attrition)
chisq.test(healthcare$Education, healthcare$Attrition)
chisq.test(healthcare$EducationField, healthcare$Attrition)
chisq.test(healthcare$EnvironmentSatisfaction, healthcare$Attrition)
chisq.test(healthcare$Gender, healthcare$Attrition)
chisq.test(healthcare$JobInvolvement, healthcare$Attrition)
chisq.test(healthcare$JobRole, healthcare$Attrition)
chisq.test(healthcare$JobSatisfaction, healthcare$Attrition)
chisq.test(healthcare$MaritalStatus, healthcare$Attrition)
chisq.test(healthcare$PerformanceRating, healthcare$Attrition)
chisq.test(healthcare$OverTime, healthcare$Attrition)
chisq.test(healthcare$RelationshipSatisfaction, healthcare$Attrition)
chisq.test(healthcare$WorkLifeBalance, healthcare$Attrition)
```
```{r}
library(knitr)
chisq_results = 
  data.frame(Variable = c("Business Travel",
                          "Department", 
                          "Education",
                          "Education Field",
                          "Environment Satisfaction",
                          "Gender",
                          "Job Involvement", 
                          "Job Role",
                          "Job Satisfaction",
                          "Marital Status",
                          "Over Time",
                          "Performance Rating",
                          "Relationship Satisfaction",
                          #"Stock Option Level",
                          "Work Life Balance"),
             Chi_Sq_Stat = c(13.59,
                             8.01, 
                             9.06, 
                             7.87,
                             23.31,
                             0.59,
                             52.00,
                             43.64,
                            11.49,
                             72.48,
                             0.11,
                             188.47,
                             1.57,
                             25.06),
             P_value = c(0.00,
                         0.01,
                         0.05, 
                         0.16,
                         0.00,
                         0.44,
                         0.00,
                         0.00,
                         0.00,
                         0.00,
                         0.73,
                         0.00,
                         0.66,
                         0.00),
             Stat_Sig = c("Yes",
                          "Yes",
                          "No",
                          "No",
                          "Yes",
                          "No",
                          "Yes",
                          "Yes",
                          "Yes",
                          "Yes",
                          "No",
                          "Yes",
                          "No",
                          "Yes"))

kable(chisq_results, 
      col.names = c("Variable","Chi-Square Statistic", "p-value", "Statistically Significant"))
```

# ANOVA for Feature Selection


```{r}
library(stats)

# ANOVA: Distance From Home and Attrition 
aov.res = aov(DistanceFromHome~Attrition, data = healthcare)
summary(aov.res)
leveneTest(aov.res)
#oneway.test(DistanceFromHome~Attrition, data = healthcare, var.equal = FALSE)



# ANOVA: Monthly Income and Attrition 
aov.res = aov(MonthlyIncome~Attrition, data = healthcare)
summary(aov.res)
leveneTest(aov.res)
#oneway.test(MonthlyIncome~Attrition, data = healthcare, var.equal = FALSE)
```



```{r}
anova_results = 
  data.frame(Variable = c("Distance From Home",
                          "Monthly Income"),
             F_Stat = c(18.87,
                        65.14),
             P_value = c(0.00,
                         0.00),
             Levene_Test = c(0.001,
                             0.00),
             Stat_Sig = c("Yes",
                          "Yes"))

kable(anova_results , 
      col.names = c("Variable","F Statistic", "p-value", "Levene's Test p-value","Statistically Significant"))
```


```{r}
healthcare2<- healthcare
```


## REGRESSION 

# GLM

```{r}
#extract all yes and nos from y
healthcare_yes <- healthcare %>% filter(Attrition=='Yes') #total 199 obs
healthcare_no <- healthcare%>% filter(Attrition=='No')

#sample 199 obs from Attrition=='No'
set.seed(200)
healthcare_no_sample<- sample_n(healthcare_no, 199)

#combine sample nos and all yes' for a balanced data set
healthcare <- rbind(healthcare_no_sample, healthcare_yes)
dim(healthcare)
```

```{r}
healthcare$Attrition<- ifelse(healthcare$Attrition =="Yes", 1, 0)  
```


```{r}
#80/20 train/test split
set.seed(200)
sample <- sample(nrow(healthcare), 0.8*nrow(healthcare), replace= F)
train<-  healthcare[sample, ]
test<- healthcare[-sample, ]
```


# ATTEMPT #1 - Fit entire model - Complex

```{r}
glm.fit1<- glm(Attrition~., data=train, family = binomial())

summary(glm.fit1)
```


```{r}
#MASS package
step(glm.fit1, direction = "both")
```


```{r}
glm.fit2<- glm(Attrition ~ Age + BusinessTravel + Department + 
    DistanceFromHome + EnvironmentSatisfaction + JobInvolvement + 
    JobRole + JobSatisfaction + MaritalStatus + NumCompaniesWorked + 
    OverTime + PercentSalaryHike + TrainingTimesLastYear + WorkLifeBalance + 
    YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager, 
    family = binomial(), data = train)

#summary(glm.fit2)

```

```{r}
#predictions saved to test df
test$GLM2Predsx <- predict(glm.fit2, test, type='response')

#covert probability predictions to yes/no format to match y (response variable)
test$GLM2Preds <- ifelse(test$GLM2Predsx>=0.5, 1 ,0 )

# Confusion matrix to compare accuracy
caret::confusionMatrix(as.factor(test$GLM2Preds), as.factor(test$Attrition))
```


# #Decision Tree Model
```{r}
library(rpart)
dt.model = rpart(Attrition~., 
                             data = train)




set.seed(200)

dt.model <- rpart(formula = Attrition~., # Y ~ all other variables in dataframe
                  data = train, # include only relevant variables
                  method = "class") # classification

rattle::fancyRpartPlot(dt.model, sub = "")
```


```{r}
dt.model$variable.importance
```


```{r}
printcp(dt.model)
```

```{r}
plotcp(dt.model)
```

```{r}
library(rpart.plot)
dt.model_2 <- prune(dt.model, 
                  cp = dt.model$cptable[which.min(dt.model$cptable[, "xerror"]), "CP"])
#rm(oj.full_class)
rpart.plot(dt.model_2, yesno = TRUE)
```


```{r}
dt.pred <- predict(dt.model_2, test, type = "class")
plot(as.factor(test$Attrition), dt.pred, 
     main = "Simple Classification: Predicted vs. Actual",
     xlab = "Actual",
     ylab = "Predicted")
```


```{r}
caret::confusionMatrix(dt.pred, as.factor(test$Attrition))
```




# RANDOM FOREST 


```{r}
library(randomForestSRC)
library(randomForest)

set.seed(123)

forest_attrition = randomForest( as.factor(Attrition)~., 
                     data = train,
                     mtry = 4,
                     importance = TRUE,
                     nodesize = 6,
                     ntree = 4000)


forest_attrition 


```

```{r}
forest_attrition$importance
```


```{r}
pred1_att = predict(forest_attrition,newdata = test)
```

# Accuracy Random Forest

```{r}
caret::confusionMatrix(pred1_att, as.factor(test$Attrition))
```


